### 2022-05-30 main/chapter_2ljets.tex
add cms run 2 references
grammar check; fix some warnings

There are certain plots and tables which, personally, I first seek when
meeting an \atlas\ search paper.
This chapter begins by presenting those results in brief in
Section~\ref{sec:2ljets_splash}.
The meat of this chapter (or if you prefer, its gory detail) follows in
Section~\ref{sec:2ljets_context} onwards.

@@

In a large number orthogonal signal regions, no significant excess is observed
above \emph{post-fit} backgrounds.
Indeed there are some significant deficits --- data below the background noise.

@@

We use two signal models in the $\twoljets$-electroweak analysis.
One is an MSSM chargino-neutralino model (C1N2) which acts to produce
$\chargino_1\textrm{--}\neutralino_2$ pairs, which decay via
$W\rightarrow q\bar q$ and
$Z\rightarrow \ell^\pm \ell^\mp$ respectively..
The other is a GMSB model which, after a variety of initial states produces
$\neutralino_1\textrm{--}\neutralino_1$ pairs.
These neutralinos decay,
via Higgs and $Z$ bosons in variable branching fractions,
to final states again containing $\ell^\pm \ell^\mp$ and jet-seeding partons.
Diagrams illustrating these C1N2 and GMSB signal models are shown in
Figure~\ref{fig:2ljets_signal_diagrams}.

Effects of some of these these signals in example signal regions of the
$\twoljets$-electroweak analysis are displayed in
Figure~\ref{fig:2ljets_signal_examples}.

@@

\begin{figure}[tp]
\centering
\begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/2ljets_sr_high_8_met_sig.pdf}
    \caption{SR-High-8}
\end{subfigure}
\hfill
\begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/2ljets_sr_llbb_mbb.pdf}
    \caption{SR-$\ell\ell bb$}
\end{subfigure}
\\[0.5em]
\begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/2ljets_sr_int_met_sig.pdf}
    \caption{SR-Int}
\end{subfigure}
\hfill
\begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/2ljets_sr_low_met_sig.pdf}
    \caption{SR-Low}
\end{subfigure}
\caption{%
Example signal region histograms with benchmark signal sample yields overlaid
as dotted lines~\cite{atlas2022searches}.
Physically, signal yields would add on top of the backgrounds.
Data are shown; regions in the the top two plots observe zero data, and
Poisson error bars for $0$ events are hidden for aesthetic reasons.
}
\label{fig:2ljets_signal_examples}
\end{figure}

@@

\begin{figure}[tp]
\centering
\includegraphics[width=0.99\textwidth]{figures/2ljets_contours_c1n2.pdf}
\caption{%
Contours for the C1N2 model in the $\twoljets$-electroweak
analysis~\cite{atlas2022searches}.
\\[0.5em]
Space below the solid red line is labelled as excluded and its dotted
neighbours show the result if all signal cross-sections are varied up and down
by theoretical error bars.
\\[0.5em]
The yellow band shows the $\pm1$-sigma region of exclusion contours
from asymptotic approximations to the prior distribution of the test statistic.
Grey areas are observed limits from the two-lepton parts
of~\cite{atlas_23l_SUSY_2016_24} and~\cite{atlas_2l_SUSY_2013_11}.
Exclusion is defined by the $95\%$ $\mathrm{CLs}$ prescription
in asymptotic approximations.
All contours are interpolated from a sparse grid.
}
\label{fig:2ljets_contours_c1n2}
\end{figure}

\begin{figure}[tp]
\centering
\includegraphics[width=0.99\textwidth]{figures/2ljets_contours_gmsb.pdf}
\caption{%
Contours for the GMSB model in the $\twoljets$-electroweak
analysis~\cite{atlas2022searches}.
\\[0.5em]
Space below the solid red line is labelled as excluded and its dotted
neighbours show the result if all signal cross-sections are varied up and down
by theoretical error bars.
\\[0.5em]
The yellow band shows the $\pm1$-sigma region of exclusion contours
from asymptotic approximations to the prior distribution of the test statistic.
Exclusion is defined by the $95\%$ $\mathrm{CLs}$ prescription
in asymptotic approximations.
All contours are interpolated from a sparse grid.
}
\label{fig:2ljets_contours_gmsb}
\end{figure}

@@

In this subsection, I claims attributions of work on the $\twoljets$ analysis
and related \atlas\ efforts.

\begin{itemize}
\item Design, implementation and execution of the electroweak part of the analysis.
\item Service as `analysis contact' from January to October 2020.
\item Production of main data inputs for all three $\twoljets$ analyses
and the RJR $3\ell$ search~\cite{atlas_rjr_3l_SUSY_2019_09},
Production of all systematic variations for the background samples;
other group members assisted with the central background sample.
Production of all electroweak signal samples and their systematic variations.
\item \TODO{?Evaluation of Monte Carlo $Z$+jets uncertainties for RJR signal regions
which indicated they exceeded $100\%$}
\item \TODO{Various collaborations with group members on scripts to conduct
tasks which were common between the searches.}
\item Integration of the electroweak analysis with ATLAS combinations and
pMSSM scan efforts, by performing their validation studies and serializing the
analysis results.
\item Preparation of electroweak results for publication in the
paper~\cite{atlas2022searches} and the public HEPData
record~\cite{maguire2017hepdata}.
\end{itemize}
Except where specified, I produced all figures presented in this thesis.
Many plotting scripts, however, are modified from versions inherited from
many \atlas\ members past and present to whom I am grateful.

@@

\clearpage
\FloatBarrier
\section{Context}
\label{sec:2ljets_context}
% previous results on 2(/3)L from ATLAS
% other constraints on these models from ATLAS/CMS
% previous region design, motivation for this work
Previous analyses of \atlas\ data have explored similar selections of
two leptons, jets, and missing transverse momentum.
Quite a few analyses, in fact.
This project exists to supplement those results with information from the full
LHC Run~2 data-set which was completed in 2018 and contains
$139~\mathrm{fb}^{-1}$ of usable proton-proton collisions at $\sqrt{s} = 13\,\eV[T]$.
Leverage of the increased data quantity is one key aim.
Analysis quality also might be improved if our designs can take lessons from
the experiences of previous work.

@@

\begin{figure}[tp]
\centering
\includegraphics[width=0.9\textwidth]{figures/2ljets_rjr_trees.pdf}
\caption{%
Decay trees for the $\twoljets$-RJR analysis.
\emph{This figure was made by the RJR team.}~\cite{atlas2022searches}.
Both diagrams represent processes also targetted in the electroweak analysis.
(left) A fully-resolved decay tree with its particles labelled.
(right) Only the $Z\rightarrow\ell\ell$ (\underline{V}ector boson)
is resolved, but the supersymmetric system recoils off
hard QCD jets labelled ``ISR'', boosting the \underline{I}nvisible particles,
to generate visibly large $\met$ in models with small mass-splittings.
The same approach is used for SR-OffShell selections in the electroweak
analysis.
}
\label{fig:2ljets_rjr_decay_trees}
\end{figure}

@@

In this way, RJR variables get some intelligence in how parent particles are
reconstructed from their visible decay products, as well as approximate
independence from uninteresting boosts, particularly those from the emission of
QCD jets outside of the supersymmetric decay tree.

@@

\subsubsection{Former excesses}
Excess data were observed in several signal regions of the partial Run~2
search ``for chargino-neutralino production using recursive
jigsaw reconstruction in final states with two or three charged
leptons''~\cite{atlas_rjr_23l_SUSY_2017_03},
which used $36.1~\mathrm{fb}^{-1}$ of proton-proton collisions at
$\sqrt{s} = 13\,\eV[T]$.
That search reported notable excesses in four signal regions.
Of these, two required two charged leptons
($\mathrm{SR}2\ell\_\mathrm{Low}$ and $\mathrm{SR}2\ell\_\mathrm{ISR}$),
and other two required three
($\mathrm{SR}3\ell\_\mathrm{Low}$ and $\mathrm{SR}3\ell\_\mathrm{ISR}$).

@@

A more precisely framed problem with RJR variables is that they tend to select
phase space which is poorly modelled by simulations.
This is plausibly explained by the fact that  RJR variables are non-trivially
related to conventional, lab-frame variables.
Simulations have been tuned to model conventional variables well, and
`systematic' variations are evaluated and derived in conventional variables.
Modelling these features accurately does not imply that all other functions of
the data are well understood.
These effects cause Monte Carlo simulations to give predictions in RJR regions
which are inaccurate and/or carry large uncertainties.

A lack of reliable simulations motivates reliance on `data-driven' background
estimates in RJR selections, which, from my perspective, are hard to interpret
or trust.
Indeed I argue that all `data-driven' methods rely not only on data but on
hypothetical extrapolations which may not be well motivated, particularly when
studying complicated features such as RJR variables.
For these reasons, I choose to not use any RJR event variables in the the
$\twoljets$-electroweak analysis.

@@

The $\twoljets$-strong search targets strongly interacting supersymmetric
effects in cases where their decays produce pairs of charged leptons.
Like the RJR and electroweak $\twoljets$ searches, the $\twoljets$-strong
signals produce jets and invisible lightest supersymmetric particles.
Unlike the RJR and electroweak parts, not all of these leptons are produced
by $Z$ boson decays.
As illustrated in Figure~\ref{fig:2ljets_strong_signal_diagrams}, decay trees
featuring sleptons can also produce lepton pairs.
Lepton pairs produced in these slepton interactions are of the same flavor
and opposite charge, like those from $Z$ boson decays, but they have different
mass dilepton mass characteristics.

% signal models
\begin{figure}[tp]
\centering
\begin{subfigure}{0.45\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/2ljets_strong_gogo_qqqqllllN1N1_N2.pdf}
    \caption{gluino-slepton}
\end{subfigure}
\hfill
\begin{subfigure}{0.47\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/2ljets_strong_gogo_qqqqZZN1N1.pdf}
    \caption{gluino-$Z^{(*)}$}
\end{subfigure}
\\[0.5em]
\begin{subfigure}{0.45\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/2ljets_strong_sqsq_qqZZN1N1.pdf}
    \caption{squark-$Z^{(*)}$}
\end{subfigure}
\caption{%
Supersymmetric signal processes in the $\twoljets$-strong search.
\emph{Figures reproduced from the paper%
}~\cite{atlas2022searches, atlas_susy_feynman}.
\\[0.5em]
All models produce same-flavour, opposite-sign lepton pairs along with jets
and $\met$.
The gluino-slepton model (a) has a kinematic endpoint in $\mll$
given in Equation~\ref{eqn:2ljets_strong_max_mll}.
The gluino-$Z^{(*)}$ and squark-$Z^{(*)}$ models decay via $Z$ boson
resonances, which may be forced off-shell for small mass splittings
$m(\neutralino_2) - m(\neutralino_1) < m(Z)$.
}
\label{fig:2ljets_strong_signal_diagrams}
\end{figure}

@@

The largest Standard Model background to this search is dileptonic $\ttbar$,
in part because no $b$-jet veto is used.
All three signal models produce same-flavor lepton pairs, but $\ttbar$
produces same-flavor ($ee$/$\mu\mu$) and different-flavor ($e\mu$) pairs at
equal rates.
This difference in flavor symmetry is exploited to estimate backgrounds in
$\twoljets$-strong,
and its previous iterations~\cite{cms_susy_2016_strong_2l_run2_1}.
The estimate works by taking different-flavor data, weighting them by factors
to calibrate for differences in the leptons' reconstruction, and histogramming
them in same-flavour regions.
Simulations for $\ttbar$ and the smaller flavour symmetric backgrounds
$tW$, $WW$, and $\tau\tau$ are replaced by this estimate.

@@

As motivated in the previous sections, I personal tastes for how searches
should be constructed.
One taste is for simulation-driven background estimates, since they are
interpretable as representations of the Standard Model and they can exceed the
precision available in data alone.
Another is for orthogonal regions, such that their likelihoods are independent
and the information in their data can be cleanly combined.

@@

\begin{figure}[tp]
\centering
\begin{subfigure}{0.62\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/2ljets_23l_sr2inthigh.png}
    \caption{SR2-int and SR2-high}
\end{subfigure}
\\[0.5em]
\begin{subfigure}{0.52\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/2ljets_23l_contours_fig_08d.png}
    \caption{Combined exclusion}
\end{subfigure}
\hfill
\begin{subfigure}{0.46\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/2ljets_23l_chosen_regions.png}
    \caption{Chosen regions}
\end{subfigure}
\caption{%
Figures reproduced from the partial Run~2 search with $2/3\ell$
selections~\cite{atlas_23l_SUSY_2016_24, hepdata.81996}.
\\[0.5em]
(a) Histogrammed distribution of $\met$ in and below SR2-int and SR2-high with
example signal models.
SR2-int requires $\met > 150\,\eV[G]$ and SR2-high requires $\met > 250\,\eV[G]$.
The rightmost bin includes overflow.
\\[0.5em]
(b) Contours for the combined $\twoljets$ and $3\ell$ channels.
Each signal point chooses a signal region which is displayed in (c).
The grey region shows exclusion from a previous search in Run~1
data~\cite{atlas_2l_SUSY_2013_11}.
\\[0.5em]
(c) Signal regions selected for each signal point in the grid.
Two-lepton selections occupy the largest mass splittings.
The $3\ell$ regions are orthogonal and combined.
The separation between SR2-low and $3\ell$ is not trivial.
}
\label{fig:2ljets_23l_stuff}
\end{figure}

@@

Searches by \cms\ have found comparable exclusions contours for the
C1N2 and GMSB models.
\TODO{CMS papers}

\section{Events}
From its name, the $\twoljets$ analysis analyses features of events with two
charged leptons (electrons or muons --- not taus), and at least one jet.
Though not nominal, the a $\met$ requirement is also implicit from the context
of $R$-parity conserving supersymmetric models.
This section serves to describe how events are selected for the $\twoljets$
analyses, and which event variables are extracted for use in its selections.

@@

\subsection{Trigger}
\label{sec:2ljets_trigger}
There would be no event variables if not for the trigger.
When living our daily lives, we are bombarded with sensory data and must pay
attention only only those gems few of any use, to avoid exhausting our
processing capacity.
Running \atlas\ is no different.

@@

The list of available triggers in \atlas\ is designed to balance data rates
against physical interest.
It includes a variety of general purpose selections, which are used for
$\twoljets$, along with more specialized targets such as vector boson fusion
or long lived particles.
The list varies through time, so we select triggers specifically in
each data-taking period.

@@

In total, we use $14$ different two-lepton trigger selections, each of which
selects for an electron pair, a muon pair, or an emu (electron-muon pair).
To describe those triggers let $\pt^{\ell_i}$ be the transverse momentum of the
$i\mathrm{th}$ hardest lepton.

In short, all triggers require $\pt^{\ell_i}$ values to exceed thresholds in
the range $8\textrm{--}26\,\eV[G]$.
To detail all trigger requirements and their variations with data period
would be too tedious and not enlightening, but their $\pt^{\ell_i}$ requirements
are summarized here.
Please do not read them all in detail.
\begin{itemize}
\item Electron pair: requirements vary between
\begin{itemize}
\item $\pt^{e_2} > 12\,\eV[G]$ and
\item $\pt^{e_2} > 17\,\eV[G]$.
\end{itemize}
\item Muon pair: various requirements include
\begin{itemize}
\item $\pt^{\mu_2} > 10\,\eV[G]$,
\item $\pt^{\mu_2} > 14\,\eV[G]$,
\item $\pt^{\mu_1} > 18\,\eV[G]~\&~\pt^{\mu_2} > 8\,\eV[G]$, and
\item $\pt^{\mu_1} > 22\,\eV[G]~\&~\pt^{\mu_2} > 8\,\eV[G]$.
\end{itemize}
\item Emu pair: various asymmetric requirements include
\begin{itemize}
\item $\pt^{e_1} > 17\,\eV[G]~\&~\pt^{\mu_1} > 14\,\eV[G]$,
\item $\pt^{e_1} > 7\,\eV[G]~\&~\pt^{\mu_1} > 24\,\eV[G]$, and
\item $\pt^{e_1} > 26\,\eV[G]~\&~\pt^{\mu_1} > 8\,\eV[G]$.
\end{itemize}
\end{itemize}
The point of listing these is to demonstrate trigger activations are
non-trivial.
This motivates lepton selections which exceed these trigger thresholds,
so as to avoid the need to model their jagged activations and variable
efficiencies near to those thresholds~\cite{
atlas_trigger_egamma_run2,
atlas_trigger_muon_run2
}.

Quite generally throughout the data used in $\twoljets$, trigger requirements
have increased in hardness with time.

\subsubsection{Alternative triggers}
Why should we only use two-lepton triggers?
Alternatives could be one-lepton triggers, for cases where the other is not
reconstructed at trigger level, or perhaps also triggers for jets and $\met$.
The answer is that the benefits of additional triggers were assessed to not
outweigh the cost of using using them.

The benefit of an additional trigger could be in increased signal yields if
that trigger do not also harmfully increase inseparable backgrounds.
A study, \emph{which was performed by Knut Vadla}, investigated the effects
of of including one-lepton triggers.
In a signal-region-like selection requiring two leptons, two jets and
$\met > 200\,\eV[G]$, this study showed that two-lepton triggers already accept
in excess of $90\%$ of events from benchmark C1N2 models.
Therefore, any additional trigger has at most a $\approx10\%$;
improvement. Adding one-lepton triggers was found to increase efficiency by
only a few percentage points, depending on the model.
We considered that to be too little to be worthwhile.

@@

\subsection{Jets}
Like leptons, hadronic jets are identified with `baseline' and `signal'
requirements with increasingly tight requirements on quality metrics and
kinematics.

>>>>

There are certain plots and tables which, personally, I first seek when
meeting an \atlas\ search paper.
This chapter begins by presenting those results in brief in
Section~\ref{sec:2ljets_splash}.
The meat of this chapter (or, if you prefer, its gory detail) follows in
Section~\ref{sec:2ljets_context} onwards.

@@

In numerous orthogonal signal regions, no significant excess is observed
above \emph{post-fit} backgrounds.
Indeed, there are some significant deficits --- data below the background noise.

@@

We use two signal models in the $\twoljets$-electroweak analysis.
One is an MSSM chargino-neutralino model (C1N2) which acts to produce
$\chargino_1\textrm{--}\neutralino_2$ pairs, which decay via
$W\rightarrow q\bar q$ and
$Z\rightarrow \ell^\pm \ell^\mp$ respectively.
The other is a GMSB model which, from a variety of initial states, produces
$\neutralino_1\textrm{--}\neutralino_1$ pairs.
These neutralinos decay,
via Higgs and $Z$ bosons in variable branching fractions,
to final states again containing $\ell^\pm \ell^\mp$ and jet-seeding partons.
Diagrams illustrating these C1N2 and GMSB signal models are shown in
Figure~\ref{fig:2ljets_signal_diagrams}.

Effects of some of these signals in example signal regions of the
$\twoljets$-electroweak analysis are displayed in
Figure~\ref{fig:2ljets_signal_examples}.

@@

\begin{figure}[tp]
\centering
\begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/2ljets_sr_high_8_met_sig.pdf}
    \caption{SR-High-8}
\end{subfigure}
\hfill
\begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/2ljets_sr_llbb_mbb.pdf}
    \caption{SR-$\ell\ell bb$}
\end{subfigure}
\\[0.5em]
\begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/2ljets_sr_int_met_sig.pdf}
    \caption{SR-Int}
\end{subfigure}
\hfill
\begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/2ljets_sr_low_met_sig.pdf}
    \caption{SR-Low}
\end{subfigure}
\caption{%
Example signal region histograms with benchmark signal sample yields overlaid
as dotted lines~\cite{atlas2022searches}.
Physically, signal yields would add on top of the backgrounds.
Data are shown; regions in the top two plots observe zero data, and
Poisson error bars for $0$ events are hidden for aesthetic reasons.
}
\label{fig:2ljets_signal_examples}
\end{figure}

@@

\begin{figure}[tp]
\centering
\includegraphics[width=0.99\textwidth]{figures/2ljets_contours_c1n2.pdf}
\caption{%
Contours for the C1N2 model in the $\twoljets$-electroweak
analysis~\cite{atlas2022searches}.
\\[0.5em]
Space below the solid red line is labelled as excluded, and its dotted
neighbours show the result if all signal cross-sections are varied up and down
by theoretical error bars.
\\[0.5em]
The yellow band shows the $\pm1$-sigma region of exclusion contours
from asymptotic approximations to the prior distribution of the test statistic.
Grey areas are observed limits from the two-lepton parts
of~\cite{atlas_23l_SUSY_2016_24} and~\cite{atlas_2l_SUSY_2013_11}.
Exclusion is defined by the $95\%$ $\mathrm{CLs}$ prescription
in asymptotic approximations.
All contours are interpolated from a sparse grid.
}
\label{fig:2ljets_contours_c1n2}
\end{figure}

\begin{figure}[tp]
\centering
\includegraphics[width=0.99\textwidth]{figures/2ljets_contours_gmsb.pdf}
\caption{%
Contours for the GMSB model in the $\twoljets$-electroweak
analysis~\cite{atlas2022searches}.
\\[0.5em]
Space below the solid red line is labelled as excluded, and its dotted
neighbours show the result if all signal cross-sections are varied up and down
by theoretical error bars.
\\[0.5em]
The yellow band shows the $\pm1$-sigma region of exclusion contours
from asymptotic approximations to the prior distribution of the test statistic.
Exclusion is defined by the $95\%$ $\mathrm{CLs}$ prescription
in asymptotic approximations.
All contours are interpolated from a sparse grid.
}
\label{fig:2ljets_contours_gmsb}
\end{figure}

@@

In this subsection, I claim attributions of work on the $\twoljets$ analysis
and related \atlas\ efforts.

\begin{itemize}
\item Design, implementation, and execution of the electroweak part of the analysis.
\item Service as `analysis contact' from January to October 2020.
\item Production of main data inputs for all three $\twoljets$ analyses
and the RJR $3\ell$ search~\cite{atlas_rjr_3l_SUSY_2019_09},
Production of all systematic variations for the background samples;
other group members assisted with the central background sample.
Production of all electroweak signal samples and their systematic variations.
\item \TODO{?Evaluation of Monte Carlo $Z$+jets uncertainties for RJR signal regions
which indicated they exceeded $100\%$}
\item \TODO{Various collaborations with group members on scripts to conduct
tasks which were common between the searches.}
\item Integration of the electroweak analysis with ATLAS combinations and
pMSSM scan efforts, by performing their validation studies and serializing the
analysis results.
\item Preparation of electroweak results for publication in the
paper~\cite{atlas2022searches} and the public HEPData
record~\cite{maguire2017hepdata}.
\end{itemize}
Except where specified, I produced all figures presented in this thesis.
Many plotting scripts, however, are modified from versions inherited from
many \atlas\ members past and present, to whom I am grateful.

@@

\clearpage
\FloatBarrier
\section{Context}
\label{sec:2ljets_context}
% previous results on 2(/3)L from ATLAS
% other constraints on these models from ATLAS/CMS
% previous region design, motivation for this work
Previous analyses of \atlas\ data have explored similar selections of
two leptons, jets, and missing transverse momentum.
Quite a few analyses, in fact.
This project exists to supplement those results with information from the full
LHC Run~2 data-set, which was completed in 2018 and contains
$139~\mathrm{fb}^{-1}$ of usable proton-proton collisions at $\sqrt{s} = 13\,\eV[T]$.
Leverage of the increased data quantity is one key aim.
Analysis quality also might be improved if our designs can take lessons from
the experiences of previous work.

@@

\begin{figure}[tp]
\centering
\includegraphics[width=0.9\textwidth]{figures/2ljets_rjr_trees.pdf}
\caption{%
Decay trees for the $\twoljets$-RJR analysis.
\emph{This figure was made by the RJR team.}~\cite{atlas2022searches}.
Both diagrams represent processes which are targetted in the electroweak
analysis.
(left) A fully-resolved decay tree with its particles labelled.
(right) Only the $Z\rightarrow\ell\ell$ (\underline{V}ector boson)
is resolved, but the supersymmetric system recoils off
hard QCD jets labelled ``ISR'', boosting the \underline{I}nvisible particles,
to generate visibly large $\met$ in models with small mass-splittings.
The same approach is used for SR-OffShell selections in the electroweak
analysis.
}
\label{fig:2ljets_rjr_decay_trees}
\end{figure}

@@

In this way, RJR variables get some intelligence in how parent particles are
reconstructed from their visible decay products, as well as approximate
independence from uninteresting boosts, particularly those from the emission of
QCD jets outside the supersymmetric decay tree.

@@

\subsubsection{Former excesses}
Excess data were observed in several signal regions of the partial Run~2
search ``for chargino-neutralino production using recursive
jigsaw reconstruction in final states with two or three charged
leptons''~\cite{atlas_rjr_23l_SUSY_2017_03},
which used $36.1~\mathrm{fb}^{-1}$ of proton-proton collisions at
$\sqrt{s} = 13\,\eV[T]$.
That search reported notable excesses in four signal regions.
Of these, two required two charged leptons
($\mathrm{SR}2\ell\_\mathrm{Low}$ and $\mathrm{SR}2\ell\_\mathrm{ISR}$),
and the other two required three
($\mathrm{SR}3\ell\_\mathrm{Low}$ and $\mathrm{SR}3\ell\_\mathrm{ISR}$).

@@

A more precisely framed problem with RJR variables is that they tend to select
phase space, which is poorly modelled by simulations.
This is plausibly explained by the fact that  RJR variables are non-trivially
related to conventional, lab-frame variables.
Simulations have been tuned to model conventional variables well, and
`systematic' variations are evaluated and derived in conventional variables.
Modelling these features accurately does not imply that all other functions of
the data are well understood.
These effects cause Monte Carlo simulations to give predictions in RJR regions
which are inaccurate and/or carry large uncertainties.

A lack of reliable simulations motivates reliance on `data-driven' background
estimates in RJR selections, which, from my perspective, are hard to interpret
or trust.
Indeed, I argue that all `data-driven' methods rely not only on data but on
hypothetical extrapolations which may not be well motivated, particularly when
studying complicated features such as RJR variables.
For these reasons, I choose to not use any RJR event variables in the
$\twoljets$-electroweak analysis.

@@

The $\twoljets$-strong search targets strongly interacting supersymmetric
effects in cases where their decays produce pairs of charged leptons.
Like the RJR and electroweak $\twoljets$ searches, the $\twoljets$-strong
signals produce jets and invisible lightest supersymmetric particles.
Unlike the RJR and electroweak parts, not all of these leptons are produced
by $Z$ boson decays.
As illustrated in Figure~\ref{fig:2ljets_strong_signal_diagrams}, decay trees
featuring sleptons can also produce lepton pairs.
Lepton pairs produced in these slepton interactions are of the same flavour
and opposite charge, like those from $Z$ boson decays, but they have different
mass dilepton mass characteristics.

% signal models
\begin{figure}[tp]
\centering
\begin{subfigure}{0.45\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/2ljets_strong_gogo_qqqqllllN1N1_N2.pdf}
    \caption{gluino-slepton}
\end{subfigure}
\hfill
\begin{subfigure}{0.47\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/2ljets_strong_gogo_qqqqZZN1N1.pdf}
    \caption{gluino-$Z^{(*)}$}
\end{subfigure}
\\[0.5em]
\begin{subfigure}{0.45\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/2ljets_strong_sqsq_qqZZN1N1.pdf}
    \caption{squark-$Z^{(*)}$}
\end{subfigure}
\caption{%
Supersymmetric signal processes in the $\twoljets$-strong search.
\emph{Figures reproduced from the paper%
}~\cite{atlas2022searches, atlas_susy_feynman}.
\\[0.5em]
All models produce same-flavour, opposite-sign lepton pairs along with jets
and $\met$.
The gluino-slepton model (a) has a kinematic endpoint in $\mll$
given in Equation~\ref{eqn:2ljets_strong_max_mll}.
The gluino-$Z^{(*)}$ and squark-$Z^{(*)}$ models decay via $Z$ boson
resonances, which may be forced off-shell for small mass-splittings
$m(\neutralino_2) - m(\neutralino_1) < m(Z)$.
}
\label{fig:2ljets_strong_signal_diagrams}
\end{figure}

@@

The largest Standard Model background to this search is dileptonic $\ttbar$,
in part because no $b$-jet veto is used.
All three signal models produce same-flavour lepton pairs, but $\ttbar$
produces same-flavour ($ee$/$\mu\mu$) and different-flavour ($e\mu$) pairs at
equal rates.
This difference in flavour symmetry is exploited to estimate backgrounds in
$\twoljets$-strong,
and its previous iterations~\cite{cms_susy_2016_strong_2l_run2_1}.
The estimate works by taking different-flavour data, weighting them by factors
to calibrate for differences in the leptons' reconstruction, and histogramming
them in same-flavour regions.
Simulations for $\ttbar$ and the smaller flavour symmetric backgrounds
$tW$, $WW$, and $\tau\tau$ are replaced by this estimate.

@@

As motivated in the previous sections, I have personal tastes for how searches
should be constructed.
One taste is for simulation-driven background estimates, since they are
interpretable as representations of the Standard Model, and they can exceed the
precision available in data alone.
Another is for orthogonal regions, such that their likelihoods are independent
and the information in their data can be cleanly combined.

@@

\begin{figure}[tp]
\centering
\begin{subfigure}{0.6\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/2ljets_23l_sr2inthigh.png}
    \caption{SR2-int and SR2-high}
\end{subfigure}
\\[0.5em]
\begin{subfigure}{0.52\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/2ljets_23l_contours_fig_08d.png}
    \caption{Combined exclusion}
\end{subfigure}
% \hfill
\begin{subfigure}{0.46\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/2ljets_23l_chosen_regions.png}
    \caption{Chosen regions}
\end{subfigure}
\caption{%
Figures reproduced from the partial Run~2 search with $2/3\ell$
selections~\cite{atlas_23l_SUSY_2016_24, hepdata.81996}.
\\[0.5em]
(a) Histogrammed distribution of $\met$ in and below SR2-int and SR2-high with
example signal models.
SR2-int requires $\met > 150\,\eV[G]$ and SR2-high requires $\met > 250\,\eV[G]$.
\\[0.5em]
(b) Contours for the combined $\twoljets$ and $3\ell$ channels.
Each signal point chooses a signal region, which is displayed in (c).
The grey region shows exclusion from a previous search in Run~1
data~\cite{atlas_2l_SUSY_2013_11}.
\\[0.5em]
(c) Signal regions selected for each signal point in the grid.
Two-lepton selections occupy the largest mass-splittings.
The $3\ell$ regions are orthogonal and combined.
}
\label{fig:2ljets_23l_stuff}
\end{figure}

@@

\section{Events}
From its name, the $\twoljets$ analysis analyses features of events with two
charged leptons (electrons or muons --- not taus), and at least one jet.
Though not nominal, a $\met$ requirement is also implicit from the context
of $R$-parity conserving supersymmetric models.
This section serves to describe how events are selected for the $\twoljets$
analyses, and which event variables are extracted for use in its selections.

@@

Searches have been performed in \cms\ for similar
C1N2~\cite{cms_susy_2022_c1n2} and GMSB models~\cite{
cms_susy_2018_partial_run2_combination,
cms_susy_2022_gmsb
}, with comparable results.

\subsection{Trigger}
\label{sec:2ljets_trigger}
There would be no event variables if not for the trigger.
When living our daily lives, we are bombarded with sensory data and must pay
attention only those gems few of any use, to avoid exhausting our
processing capacity.
Running \atlas\ is no different.

@@

The list of available triggers in \atlas\ is designed to balance data rates
against physical interest.
It includes a variety of general purpose selections, which are used for
$\twoljets$, along with more specialized targets such as vector boson fusion
or long-lived particles.
The list varies through time, so we select triggers specifically in
each data-taking period.

@@

In total, we use $14$ different two-lepton trigger selections, each of which
selects for an electron pair, a muon pair, or an emu (electron-muon pair).
To describe those triggers, let $\pt^{\ell_i}$ be the transverse momentum of the
$i\mathrm{th}$ hardest lepton.

In short, all triggers require $\pt^{\ell_i}$ values to exceed thresholds in
the range $8\textrm{--}26\,\eV[G]$.
To detail all trigger requirements and their variations with data period
would be too tedious and not enlightening, but their $\pt^{\ell_i}$ requirements
are summarized here.
Please do not read them all in detail.
\begin{itemize}
\item Electron pair: requirements vary between
\begin{itemize}
\item $\pt^{e_2} > 12\,\eV[G]$ and
\item $\pt^{e_2} > 17\,\eV[G]$.
\end{itemize}
\item Muon pair: various requirements include
\begin{itemize}
\item $\pt^{\mu_2} > 10\,\eV[G]$,
\item $\pt^{\mu_2} > 14\,\eV[G]$,
\item $\pt^{\mu_1} > 18\,\eV[G]~\&~\pt^{\mu_2} > 8\,\eV[G]$, and
\item $\pt^{\mu_1} > 22\,\eV[G]~\&~\pt^{\mu_2} > 8\,\eV[G]$.
\end{itemize}
\item Emu pair: various asymmetric requirements include
\begin{itemize}
\item $\pt^{e_1} > 17\,\eV[G]~\&~\pt^{\mu_1} > 14\,\eV[G]$,
\item $\pt^{e_1} > 7\,\eV[G]~\&~\pt^{\mu_1} > 24\,\eV[G]$, and
\item $\pt^{e_1} > 26\,\eV[G]~\&~\pt^{\mu_1} > 8\,\eV[G]$.
\end{itemize}
\end{itemize}
The point of listing these is to demonstrate trigger activations are
non-trivial.
This motivates lepton selections which exceed these trigger thresholds,
to avoid the need to model their jagged activations and variable
efficiencies near to those thresholds~\cite{
atlas_trigger_egamma_run2,
atlas_trigger_muon_run2
}.

Quite generally, throughout the data used in $\twoljets$, trigger requirements
have increased in hardness with time.

\subsubsection{Alternative triggers}
Why should we only use two-lepton triggers?
Alternatives could be one-lepton triggers, for cases where the other is not
reconstructed at trigger level, or perhaps also triggers for jets and $\met$.
The answer is that the benefits of additional triggers were assessed to not
outweigh the cost of using them.

The benefit of an additional trigger could be in increased signal yields if
that trigger does not also harmfully increase inseparable backgrounds.
A study, \emph{which was performed by Knut Vadla}, investigated the effects
of including one-lepton triggers.
In a signal-region-like selection requiring two leptons, two jets and
$\met > 200\,\eV[G]$, this study showed that two-lepton triggers already accept
in excess of $90\%$ of events from benchmark C1N2 models.
Therefore, any additional trigger has at most a $\approx10\%$;
improvement. Adding one-lepton triggers was found to increase efficiency by
only a few percentage points, depending on the model.
We considered that to be too little to be worthwhile.

@@

\subsection{Jets}
Like leptons, hadronic jets are identified with `baseline' and `signal'
requirements, with increasingly tight requirements on quality metrics and
kinematics.
